---
title: "homework3"
author: "Collin"
date: "4/28/2021"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggthemes)
library(leaps)
library(caret)
```


# Question 8

*a)* *Use the rnorm() function to generate a predictorXof lengthn= 100, as well as a noise vector"of lengthn=100.*
```{r}
predictorX = rnorm(100, 0,10)
noise = rnorm(100, 0, 20)



```


*b)* *generate a response vectorY of length n= 100 according to the model*
\[
Y = \beta_{0} + \beta_{1}X_{1} + \beta_{2}X_{2}^2 + \beta_{3}X_{3}^3
\]
```{r}
beta0 = 3
beta1 = 2
beta2 = -.5
beta3 = .05

Y = beta0 + beta1*predictorX + beta2*predictorX + beta3*predictorX + noise

df = data.frame(Y = Y, X= predictorX)
```



*c)* *Use theregsubsets()function to perform best subset selection in order to choose the best model containing the predictorsX,X2,...,X10. What is the best model obtained according to Cp,BIC,andadjustedR2?Showsomeplotstoprovideevidencefor your answer, and report the coefficients of the best model ob-tained. Note you will need to use the data.frame()function tocreate a single data set containing bothXandY.*
```{r}



fit = regsubsets(Y ~ poly(X, 10), data = df, nvmax = 10)

fit_summary = summary(fit)
fit_summary$cp
fit_summary$adjr2
fit_summary$bic
which.min(fit_summary$cp)


```
Based on Mallow's CP, the best model is model #4.
The coefficients are reported as follows:

```{r}
coef(fit, id = 4)


#found these plots online and thought they looked pretty cool so I included them
# tibble(Cp = fit_summary$cp,
#            BIC = fit_summary$bic,
#            AdjR2 = fit_summary$adjr2) %>%
#     mutate(id = row_number()) %>%
#     gather(value_type, value, -id) %>%
#     ggplot(aes(id, value, col = value_type)) +
#     geom_line() + geom_point() + ylab('') + xlab('Number of Variables Used') +
#     facet_wrap(~ value_type, scales = 'free') +
#     theme_minimal() + scale_x_continuous(breaks = 1:10)


```


*d)* *Repeat (c), using forward stepwise selection and also using back-wards stepwise selection. How does your answer compare to theresults in (c)?*
```{r}

#forward
forward_model = regsubsets(Y ~ poly(X, 10), data = df,
                           method = "forward")
forward_summary = summary(forward_model)
which.min(forward_summary$cp)
which.min(forward_summary$bic)
which.max(forward_summary$adjr2)
coef(forward_model, id = 4)

#best backward
backward_model = regsubsets(Y ~ poly(X, 10), data = df,
                           method = "backward")
backward_summary = summary(backward_model)
which.min(backward_summary$cp)
which.min(backward_summary$bic)
which.max(backward_summary$adjr2)
coef(forward_model, id = 4)


```
Again using Mallow's CP as my selection criteria, it appears that forward, backward, and best subset selection methods are all in agreement; that is, model #4 is found to be the "best" model in each scenario.


```{r include=FALSE}

?postResample
model_back <- train(Y ~ poly(X, 10), data = df, 
                    method = 'glmStepAIC', direction = 'backward', 
                    trace = 0,
               trControl = trainControl(method = 'none', verboseIter = FALSE))

postResample(predict(model_back, df), df$Y)

summary(model_back$finalModel)

x_poly <- poly(df$X, 10)

colnames(x_poly) <- paste0('poly', 1:10)
model_forw <- train(y = Y, x = x_poly,
                    method = 'glmStepAIC', direction = 'forward',
                    trace = 0,
               trControl = trainControl(method = 'none', verboseIter = FALSE))

postResample(predict(model_forw, data.frame(x_poly)), df$Y)
summary(model_forw$finalModel)

```



*e)* *Now fit a lasso model to the simulated data, again usingX,X2,...,X10as predictors. Use cross-validation to select the optimalvalue of lambda. Create plots of the cross-validation error as a functionofλ.Report theresultingcoefficient estimates, and discuss theresults obtained.*

```{r}
library(gamlr)
#estimate LASSO
model_matrix = model.matrix(Y ~ poly(X,10), data = df)

lasso = cv.gamlr(model_matrix, Y)
summary(lasso)


#SElection criteria:
#CV-Min: chooses smalleset out of sample SSE 
lambda_min = lasso$seg.min
lasso$gamlr$lambda[lambda_min] #lambda that minimizes SSE

#1standard error away rule
lambda_1se = lasso$seg.1se
lasso$gamlr$lambda[lambda_1se]

#Aic

#Bic

#Path plot
plot(lasso$gamlr)

plot(lasso)

coef(lasso, select = "min")

```
Using cross-validation, I find that the 1st, 2nd, and 6th terms are included in the optimal model, in addition to the intercept.



```{r include=FALSE}
#using caret
lasso_model <- train(Y ~ poly(X, 10), data = df,
                     method = 'glmnet',
                     trControl = trainControl(method = 'cv', number = 10),
                     tuneGrid = expand.grid(alpha = 1,
                                          lambda = seq(0.001, 0.2, by = 0.005)))
plot(lasso_model)
?postResample
coef(lasso_model$finalModel)
postResample(predict(lasso_model, df), df$Y) #calculates RMSE, RSquared, MAE, using predictions from the model and the actual data
?glmnet
```




*f)* *now generate a response vectorY according to the model*
```{r}
newY <- 3 + 8*predictorX^7 + noise

new_df = data_frame(Y = newY, X = predictorX)
fit <- regsubsets(Y ~ poly(X, 10), data = new_df, nvmax = 10)

fit_summary <- summary(fit)


# tibble(Cp = fit_summary$cp,
#            BIC = fit_summary$bic,
#            R2 = fit_summary$adjr2) %>%
#     mutate(id = row_number()) %>%
#     gather(value_type, value, -id) %>%
#     ggplot(aes(id, value, col = value_type)) +
#     geom_line() + geom_point() + ylab('') + xlab('Number of Variables Used') +
#     facet_wrap(~ value_type, scales = 'free') +
#     theme_minimal() + scale_x_continuous(breaks = 1:10)
```



#2

## a)
```{r}
library(ISLR)
data = College
dim(data)

n = nrow(data)
K = 5
foldid = rep(1:K, each = ceiling(n/K))[sample(1:n)]
?rep

train = data[which(foldid != 1),]
test = data[-which(foldid !=1),]



```



##b)
```{r}

ols = lm(Apps ~ ., data = train)
#predict
ols_predictions = predict(ols, test)

#calcualte out of sample SSE
ols_oos = mean( (test$Apps - ols_predictions )^2 )


#use the postResmaple function and see if you can get it
postResample(test$Apps,ols_predictions)



```


##c)
```{r message=FALSE, warning=FALSE, include=FALSE}
#Francis im so sorry I couldn't figure out why it continues to print the entire model matrix. it doesn't make any sense
library(glmnet)
#estimate ridge regression on training data

X <- model.matrix(Apps~ ., data = train)
```

```{r}
y = train$Apps

grid = 10^seq(4,-2,length = 100)

ridge = cv.glmnet(X,y, lambda = grid)

#obtain predictions and find out of sample deviance
xtest = model.matrix(Apps ~ ., data = test)
ridge_pred = predict(ridge, newx = xtest, select = ridge$lambda.min)

#calculate oos 
ridge_oos = mean( (test$Apps - ridge_pred)^2)
ridge_oos


```



##d)
```{r}
library(gamlr)
library(tidyverse)
#estimate the LASSO model on the training data
X - model.matrix(Apps ~ ., data = train)
y = train$Apps

lasso = cv.gamlr(X,y, nlambda = 1000, lambda.start = Inf, lambda.min.ratio = .001)

#obtain predictions
xtest = model.matrix(Apps ~ ., data = test)
lasso_pred = predict(lasso, newdata = xtest, select = "min")

lasso_oos = mean( (test$Apps - lasso_pred)^2)


RSS = c(ols_oos,ridge_oos,lasso_oos)
Model = c("OLS", "Ridge", "Lasso")
out = rbind(Model, round(RSS,2))
out


```
Here we see that our Lasso model results in the lowest sum of squared errors.

#3
#a)
```{r}
set.seed(10101)
#set p = 20, n = 1000
p = 20
n = 1000

X = matrix(rnorm(n*p),n,p)
#generate beta and outcome

B = rnorm(p)
B[3] = 0
B[9] = 0
B[4] = 0

e = rnorm(n)
y = X %*%B + e

#df
data = data.frame(y,X)


```





```{r}
#training data index
test_sample = sample(seq(1000),100, replace = FALSE)
sample(seq(1000),100,replace = FALSE)
test = data[test_sample,]
train = data[-test_sample,]





```


#c)
```{r}
bestsubset = regsubsets(y~.,data = train, nvmax = 20)

#graph
matrix = model.matrix(y ~ .,data = train)

bestsubset_is = rep(NA,p)
for(i in 1:20){
    coefi = coef(bestsubset,id = i)
    pred = matrix[,names(coefi)]%*%coefi
    bestsubset_is[i] = mean((train$y - pred)^2)
}
plot(bestsubset_is, ylab = "Training MSE", pch = 20, type = "b")

```
The important thing to note here that the training MSE *always* decreases with the addition of new predictors.




```{r}
matrix = model.matrix(y ~ .,data = test)

bestsubset_oos = rep(NA,p)
for(i in 1:20){
    coefi = coef(bestsubset,id = i)
    pred = matrix[,names(coefi)]%*%coefi
    bestsubset_oos[i] = mean((test$y - pred)^2)
}
plot(bestsubset_oos, ylab = "testing MSE", pch = 20, type = "b")


```
Now note that each additional predictor does not always result in a reduction of *testing* MSE. This is an incredibly important distinction.

#e)
```{r}
#in sample MSE
which.min(bestsubset_is)

#OOS
which.min(bestsubset_oos)


```
The model takes on its lowest MSE value with 15 predictors (compared to the training model's lowest MSE value which unsurprisingly comes with the maximum number of predictors we tried).







